---
title: "Beta regression"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Beta regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  cache = FALSE
)
```

We present how to perform the simulation studies in @guerrier2020general for beta regression. We warn the reader that the code presented here may take a very long time to run on their personal computer. Indeed, for the paper, the computations were performed at University of Geneva on "Baobab" and "Yggdrasil" HPC clusters. We present the regular beta regression and the beta regression with rounded responses. 

## Setting
Some of the setting's parameters are common to both situations.

```{r setting}
# packages
library(IBpaper)
library(ib)
library(betareg)

# simulation specifics
MC <- 1000 # number of simulations
n <- 2000 # sample size
p <- 300 # number of coefficients
# Non-zero coefficients:
# intercept is -0.5
beta_nz <- c(-.5,rep(1,5),rep(-1.5,5),rep(2,5))
# Coefficients:
beta <- c(beta_nz,rep(0,p-15))
gamma <- 5 # precision parameter
theta <- c(beta,gamma)

# seeds for random number generator
set.seed(895724)
seed <- vector("list",3)
seed$process <- sample.int(1e7,MC) # for the response
seed$design <- sample.int(1e7,MC) # for the design
seed$ib <- sample.int(1e7,MC) # for the iterative bootstrap

# IB specifics
H <- 200 # number of simulated estimators
bbc_control <- ibControl(H=H, maxit=1L, constraint=FALSE) # control for bootstrap bias correction (BBC)
ib_control <- ibControl(H=H, maxit=50L, constraint=FALSE) # control for iterative bootstrap (IB)
```


## Regular beta regression
Here is the code for the simulation of beta regression:

```{r regular beta}
# For saving the results
res <- list(mle = matrix(ncol=p+2,nrow=MC),
            bbc = matrix(ncol=p+2,nrow=MC),
            jini = matrix(ncol=p+2,nrow=MC))

for(m in 1:MC){
  ##------- simulate the process ---------
  set.seed(seed$process[m]) # set the seed
  x <- matrix(rnorm(n*p, sd = p^(-.5)), nrow = n) # simulate the design
  betareg_object <- make_betareg(x, theta) # see ?make_betareg
  # simulate beta responses
  y <- simulation(betareg_object, 
                  control = list(seed=seed$process[m])) 
  
  ##------ MLE estimation ----------------
  # Note: here the MLE is the initial estimator
  fit_mle <- NULL
  try(fit_mle <- betareg(y ~ x), silent=T)
  if(is.null(fit_mle)) next
  res$mle[m,] <- coef(fit_mle)
  
  ##------ Bootstrap bias correction -----
  bbc_control$seed <- seed$ib[m] # update the seed
  fit_bbc <- ib(fit_mle, control = bbc_control) # compute BBC , see ? ib::ib for more details
  res$bbc[m,] <- getEst(fit_bbc) # retrieve estimator
  
  ##------ IB bias correction ------------
  ib_control$seed <- seed$ib[m] # update the seed
  fit_jini <- ib(fit_mle, control = ib_control) # compute IB
  res$jini[m,] <- getEst(fit_jini) # retrieve estimator
  # getIteration(fit_jini), if one wants to retrieve the number of iterations

  # print the iteration
  cat(m,"\n")
}
```

## Beta regression with rounded responses
We need first to specify the "rounding process". For that, we need to specify how to generate rounded responses:
```{r interference}
# Function to simulate responses with rounding
# see ?ib::simulation and ?ib::control for more details 
rounding <- function(object, control, extra=NULL){
  simulate_betareg <- getFromNamespace("simulate_betareg", ns = "ib")
  y <- simulate_betareg(object, control$H)
  n <- length(y) / control$H
  y <- round(y,1)
  y <- (y*(n-1) + 0.5)/n
  matrix(y, ncol=control$H)
}
ib_control <- ibControl(H = H, maxit = 50L, sim = rounding) # control for iterative bootstrap (IB)
```

```{r beta regression with rounded responses}
# For saving the results
res <- list(mle = matrix(ncol=p+2,nrow=MC),
            initial = matrix(ncol=p+2,nrow=MC),
            jini = matrix(ncol=p+2,nrow=MC))

for(m in 1:MC){
  ##------- simulate the process ---------
  set.seed(seed$process[m]) # set the seed
  x <- matrix(rnorm(n*p, sd = p^(-.5)), nrow = n) # simulate the design
  betareg_object <- make_betareg(x, theta) # see ?make_betareg
  # simulate beta regression with rounded responses
  y <- simulation(betareg_object, 
                  control = list(seed = seed$process[m], 
                                 sim = rounding)) 
  
  ##------ Initial estimator (inconsistent) ----------------
  fit_initial <- NULL
  try(fit_initial <- betareg(y ~ x), silent=T)
  if(is.null(fit_initial)) next
  res$mle[m,] <- coef(fit_initial)
  
  ##------ MLE estimation ----------------
  fit_mle <- NULL
  sv <- fit_initial # starting values
  sv[p+2] <- log(sv[p+2]) # log transform
  try(fit_mle <- em(y,cbind(1,x),sv), silent=T) # see ?em
  if(!is.null(fit_em)){
    res$mle[m,] <- fit_mle$par
    res$mle[m,p+2] <- exp(fit_mle$par[p+2])
  }
  
  ##------ IB bias correction ------------
  ib_control$seed <- seed$ib[m] # update the seed
  fit_jini <- ib(fit_initial, control = ib_control) # compute IB
  res$jini[m,] <- getEst(fit_jini) # retrieve estimator
  # getIteration(fit_jini), if one wants to retrieve the number of iterations

  # print the iteration
  cat(m,"\n")
}
```

## References
