---
title: "Negative Binomial Regression"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Negative binomial regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  cache = FALSE
)
```

Here we present how to perform the simulation studies in @zhang2022flexible for the negative binomial regression, with and without interference. We warn the readers that the code presented here may take time to run on a personal computer. Indeed, for the paper, the computations were performed on the "Baobab" and "Yggdrasil" HPC clusters provided by the University of Geneva. 

## Setting

We use the following setting for the negative binomial regression, with and without interference.

```{r setting}
# packages
library(IBpaper)
library(ib)
library(MASS)

# simulation specifics
MC <- 1000 # number of simulations
n <- 200 # sample size
p <- 40 # number of coefficients
# Non-zero coefficients:
# intercept is 2
beta_nz <- c(2.0,1.0,-1.0)
# Coefficients:
beta <- c(beta_nz,rep(0,p-2))
alpha <- 0.7 # overdispersion parameter

# seeds for random number generator
set.seed(895724)
seed <- vector("list",3)
seed$process <- sample.int(1e7,MC) # for the response
seed$design <- sample.int(1e7,MC) # for the design
seed$ib <- sample.int(1e7,MC) # for the iterative bootstrap

# IB specifics
H <- 200 # number of simulated estimators
bbc_control <- ibControl(H=H, maxit=1L) # control for bootstrap bias correction (BBC)
ib_control <- ibControl(H=H, maxit=50L) # control for iterative bootstrap (IB)
```


## Negative binomial regression without interference

Here is the code for the simulation of the negative binomial regression computed on the actual responses without interference. 

```{r regular negative binomial}
# For saving the results
res <- list(mle = matrix(ncol=p+2,nrow=MC),
            bbc = matrix(ncol=p+2,nrow=MC),
            jini = matrix(ncol=p+2,nrow=MC))

for(m in 1:MC){
  ##------- simulate the process ---------
  set.seed(seed$process[m]) # set the seed
  x <- matrix(rnorm(n*p, sd = 2 * sqrt(2) / sqrt(p)), nrow = n) # simulate the design
  negbin_object <- make_negbin(x, beta, 1 / alpha) # see ?make_negbin
  # simulate negative binomial responses
  y <- simulation(negbin_object, 
                  control = list(seed=seed$process[m])) 
  
  ##------ MLE estimation ----------------
  # Note: here the MLE is the initial estimator
  fit_mle <- NULL
  try(fit_mle <- glm.nb(y ~ x), silent=T)
  if(is.null(fit_mle)) next
  res$mle[m,1:(p+1)] <- coef(fit_mle)
  res$mle[m,p+2] <- 1 / fit_mle$theta
  
  ##------ Bootstrap bias correction -----
  bbc_control$seed <- seed$ib[m] # update the seed
  fit_bbc <- ib(fit_mle, control = bbc_control, 
                extra_param = TRUE) # compute BBC , see ? ib::ib for more details
  res$bbc[m,] <- getEst(fit_bbc) # retrieve estimator
  
  ##------ IB bias correction ------------
  ib_control$seed <- seed$ib[m] # update the seed
  fit_jini <- ib(fit_mle, control = ib_control, extra_param = TRUE) # compute IB
  res$jini[m,] <- getEst(fit_jini) # retrieve estimator
  # getIteration(fit_jini), if one wants to retrieve the number of iterations

  # print the iteration
  cat(m,"\n")
}
```

## Negative binomial regression with interfered responses

We first need to specify the interference mechanism, i.e. how the responses are interfered. 

```{r interference}
# Function to simulate interfered responses
# see ?ib::simulation and ?ib::control for more details 
random_censoring <- function(object, control, extra=NULL){
  simulate_negbin <- getFromNamespace("simulate_negbin", ns = "ib")
  y <- simulate_negbin(object, control$H)
  u <- rpois(length(y), lambda = 3)
  y[y<=u] <- u[y<=u]
  matrix(y, ncol=control$H)
}
ib_control <- ibControl(H = H, maxit = 50L, sim = random_censoring) # control for iterative bootstrap (IB)
```

Here is the code for the simulation of the negative binomial regression computed on the interfered responses. 

```{r negative binomial regression with interfered responses}
# For saving the results
res <- list(mle = matrix(ncol=p+2,nrow=MC),
            initial = matrix(ncol=p+2,nrow=MC),
            jini = matrix(ncol=p+2,nrow=MC))

for(m in 1:MC){
  ##------- simulate the process ---------
  set.seed(seed$process[m]) # set the seed
  x <- matrix(rnorm(n*p, sd = 2 * sqrt(2) / sqrt(p)), nr=n) # simulate the design
  negbin_object <- make_negbin(x, beta, 1 / alpha)# see ?make_negbin
  # simulate negative binomial with interfered responses
  y <- simulation(negbin_object, 
                  control = list(sim=random_censoring,
                                 seed=seed$process[m])) 
  
  ##------ MLE estimation ----------------
  fit_mle <- NULL
  try(fit_mle <- glm_negbin(y, x, lambda = 3), silent = TRUE) # see ? glm_negbin
  if(!is.null(fit_em)){
    res$mle[m,] <- fit_mle$par
  }
  
  ##------ Initial estimator (inconsistent) ----------------
  fit_initial <- NULL
  try(fit_initial <- glm.nb(y ~ x), silent=T)
  if(is.null(fit_initial)) next
  res$initial[m,1:(p+1)] <- coef(fit_initial)
  res$initial[m,p+2] <- 1 / fit_initial$theta
  
  ##------ IB bias correction ------------
  ib_control$seed <- seed$ib[m] # update the seed
  fit_jini <- ib(fit_initial, control = ib_control, extra_param = TRUE) # compute IB
  res$jini[m,] <- getEst(fit_jini) # retrieve estimator
  # getIteration(fit_jini), if one wants to retrieve the number of iterations

  # print the iteration
  cat(m,"\n")
}
```

## References
