---
title: "Alcohol Consumption Data Analysis"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Alcohol Consumption Data Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  cache = FALSE
)
```

Here we present the analysis on the alcohol consumption real data in @zhang2022flexible. The data is made available in the package, see `?data(student)`. This dataset was originally collected and studied by @cortez2008using. In this analysis, we suppose that the binary response `alc` is subject to a false negative rate of 5\%. We warn the readers that the code presented here may take time to run on a personal computer. Indeed, for the paper, the computations were performed on the "Baobab" and "Yggdrasil" HPC clusters provided by the University of Geneva. 
 

## Setting

```{r setting}
# packages
library(IBpaper)
library(ib)

# for simulating logistic with misclassified response
misclassification <- function(object, control, extra){
  ftd <- fitted(object)
  n <- length(ftd)
  N <- n * control$H
  u1 <- 0.0
  u2 <- 0.05
  mu_star <- u1 * (1 - rep(ftd,control$H)) + (1 - u2) * rep(ftd,control$H)
  matrix(ifelse(runif(N) > mu_star,0,1), ncol=control$H)
}

# Verify where a value lies given a confidence interval
check_ci <- function(ci,theta){
  if(!is.matrix(ci)) ci <- as.matrix(ci)
  if(ncol(ci)!=2) ci <- t(ci)
  if(nrow(ci)!=length(theta)) stop("verify ci and theta dimensions match")
  apply(cbind(ci,theta), 1, function(x){
    if(x[3] <= min(x[1:2])){"left"}
    else if(x[3] >= max(x[1:2])){"right"}else{"center"}
  })
}

# IB specifics
H <- 200 # number of simulated estimators
ib_control <- ibControl(H = H, sim = misclassification, maxit = 100L) # control for iterative bootstrap (IB)
B <- 200 # number of bootstrap replicates
set.seed(432)
seeds <- sample.int(1e7,B) # seeds for bootstrap

# For coverage
alpha <- 0.05 # nominal level for confidence interval
```

## Real data analysis

The code for the analysis on the real data is provided as follows.

```{r real data}
## Fit initial estimator
fit_initial <- glm(alc ~ ., family = binomial(link = "logit"), 
                   data = student)

## Fit the MLE
x <- as.matrix(student[,-1])
y <- student[,1]
fit_mle <- logistic_misclassification_mle(x, y, fp = 0, fn = 0.05)

## Fit the JINI
fit_ib <- ib(fit_initial, control = ib_control)

## Retrieve coefficients
beta_hat <- fit_mle
beta_tilde <- coef(fit_ib)
p <- length(beta_hat)

## Standard errors for the MLE (based on Fisher information)
se1 <- sqrt(diag(inverse_FIM(x,beta_hat,0.0,0.05)))

## Asymptotic confidence interval for MLE
ci_mle <- beta_hat + se1 * matrix(rep(qnorm(c(alpha/2,1-alpha/2)),p),nc=2,byr=T)

## Standard errors for the JINI (estimated via bootstrap)
logistic_object <- make_logistic(x, beta_tilde)
boot <- matrix(nrow = B, ncol = p)
for(b in 1:B){
  ##------- simulate the process ---------
  # simulate logistic
  y <- simulation(logistic_object, control=list(seed=b, sim=misclassification))

  ##------ Initial estimation ----------------
  fit_tmp <- glm(y ~ x, family=binomial(link="logit"))

  ##------ IB bias correction ------------
  ib_control$seed <- seeds[b] # update seed in control
  fit_ib <- ib(fit_tmp, control = ib_control)
  boot[b,] <- getEst(fit_ib)
}

se2 <- sqrt(diag(cov(boot)))

## Asymptotic confidence interval for JINI
ci_jini <- beta_tilde + se2 * matrix(rep(qnorm(c(alpha/2,1-alpha/2)),p),nc=2,byr=T)
```


## Numerical experiment

For the numerical experiment, we use `beta_tilde` from the real data analysis as the true parameter.

```{r numerical study}
MC <- 1e4 # number of simulation
logistic_object <- make_logistic(x, beta_tilde) # for simulating response
# seed for RNG
set.seed(5847329)
seed <- vector("list",3)
seed$process <- sample.int(1e7,MC)
seed$ib <- sample.int(1e7,MC)
seed$boot <- sample.int(1e7,B)

# for saving results
res <- list(
  mle = matrix(nrow = MC, ncol = p),
  lci_mle = matrix(nrow = MC, ncol = p),
  coverage_mle = matrix(nrow = MC, ncol = p),
  ib = matrix(nrow = MC, ncol = p),
  lci_ib = matrix(nrow = MC, ncol = p),
  coverage_ib = matrix(nrow = MC, ncol = p)
)

ib_control2 <- ib_control # copy for bootstrap

# Simulations (careful, it is very long)
for(m in 1:MC){
  ##------- simulate the process ---------
  # simulate logistic
  y <- simulation(logistic_object, 
                  control = list(seed = seed$process[m], 
                                 sim = misclassification))
  
  ##------ Initial estimation ----------------
  fit_initial <- glm(y ~ x, family=binomial(link="logit"))
  
  ##------ MLE estimation ----------------
  fit_mle <- logistic_misclassification_mle(x,y,0.0,0.05)
  beta_hat <- fit_mle
  ci_mle <- beta_hat - sqrt(diag(inverse_FIM(x,beta_hat,0.0,0.05))) * matrix(rep(qnorm(alpha[2:1]),p),nc=2,byr=T) 
  res$mle[m,] <- beta_hat
  res$lci_mle[m,] <- apply(ci_mle,1,diff)
  res$coverage_mle[m,] <- check_ci(ci_mle,beta)
  
  ##------ IB bias correction ------------
  ib_control$seed <- seed$ib[m]
  fit_ib <- ib(fit_initial, control = ib_control)
  beta_tilde <- getEst(fit_ib)
  
  # bootstrap approximation of covariance
  lo <- make_logistic(x, beta_tilde)
  boot <- matrix(nrow = B, ncol = p)
  for(b in 1:B){
    ##------- simulate the process ---------
    # simulate logistic
    y <- simulation(lo, control=list(seed=b, sim=misclassification))
    
    ##------ MLE estimation ----------------
    fit_tmp <- glm(y~x, family=binomial(link="logit"))
    
    ##------ IB bias correction ------------
    ib_control2$seed <- seed$boot[b]
    fit_ib <- ib(fit_tmp, control = ib_control2)
    boot[b,] <- getEst(fit_ib)
  }
  se2 <- sqrt(diag(cov(boot)))
  ci_ib <- beta_tilde - se2 * matrix(rep(qnorm(alpha[2:1]),p),nc=2,byr=T) 
  res$ib[m,] <- beta_tilde
  res$lci_ib[m,] <- apply(ci_ib,1,diff)
  res$coverage_ib[m,] <- check_ci(ci_ib,beta)
  
  cat(m,"\n")
}
```



## References
